---
title: "Notebook 2: Part 2"
output: html_notebook
---

# Health Burden

In this first code section, load your data and any necessary libraries (`dplyr`, `tidyr`, `ggplot`, any others you like).
```{r, echo=F}
# Set up (install packages if you don't have them)
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
risk.data <- read.csv('./data/prepped/risk-data.csv', stringsAsFactors = FALSE) 
```

## Data Structure
To get a basic sense of your dataset, check the following:

- What is the size of the dataset (rows, columns)?
- How many countries are present in the dataset?
- What are the age-groups in the dataset?
- What are the variables (columns) present in the dataset?
- What is the data type of each variable?
- What are the summary statistics of each quantitative variable (_hint: `summary()`_)

```{r}
row_count = nrow(risk.data)
col_count = ncol(risk.data)
number_of_countries = length(unique(risk.data$country))
age_groups = unique(risk.data$age)
variables = colnames(risk.data)
variable_types = sapply(risk.data, typeof)
summary_stats = summary(risk.data)

print(row_count)
print(col_count)
print(number_of_countries)
print(age_groups)
print(variables)
print(variable_types)
print(summary_stats)
```

## Univariate Analysis
To explore the distribution of each risk-column in a visual way, create a histogram of each risk-column of interest (e.g., `alcohol_use`, `drug_use`, etc.). For easier comparisons, you may want to store each histogram in a variable, then render all of the charts using the `grid.arrange` function from the `gridExtra` package.

If you want to do this in an more advanced (and concise!) `R` way, you should **reshape your data** using the `gather` function to combine the risks (which are currently each their own column) into two columns (one for the _name_ of the risk, and the other for the _value_ of the death rate of that risk). Then, you can use the `facet_wrap` method of the ggplot library.

- Note what you find surprising!


```{r}
# Make histograms for each risk column
data.reshape = gather(risk.data, "name", "value", alcohol_use:smoking)
ggplot(data.reshape, aes(x = value)) + geom_histogram() +  facet_wrap(~name)


```

_list some observations from your histograms here using Markdown code_:

- Meat consumption does not play a role in death rates.
- I would've expected drug use to also play a role in death rates but this data shows that it doesn't as much in comparison to low exercise.
- Low exercise and smoking use are relatively at the same risk

## Univariate analysis (by age)
In this section, you should **write a function** to investigate how a given risk (column) varies by **age group** (the risk -- column name -- will be the argument to your function). More specifically, the function should create a set of histograms (one for each age-group) to show the distribution of death rates for a given risk factor. You should use the `facet_wrap` function (from `ggplot2`) to create the multi-plot layout.

```{r}
# Write your function here
by_age <- function(risk_name) {
  factored.risk <- select(risk.data, age, risk_name)
  ggplot(factored.risk, aes_string(x = risk_name)) + geom_histogram() +  facet_wrap(~age)
}

# Create a plot to demonstrate that it works
by_age('alcohol_use')

```

_list some observations from your histograms here using Markdown code_:

- There are a lot of deaths between the 15-49 age range for alcohol use. 
- For most risks, the 70+ age range seems to have the same value for death rates.


## Univariate analysis (by sex)
In this section, you should investigate how a given risk (column) varies by **sex**. To do this, **write a function** to create a single histogram with _overlapping distributions_ for males and females (for a given risk factor). 

_Hint_: set the `fill` of your `ggplot()` aesthetic, and turn down the _alpha_ (opacity) to help reduce occlusion in the chart. 

```{r}
# Write your function here
by_sex <- function(risk_name) {
  df = select(risk.data, sex, risk_name)
  ggplot(df,aes_string(x=risk_name)) + 
    geom_histogram(data=subset(df,sex == 'male'), aes(fill=sex), alpha = 0.6) +
    geom_histogram(data=subset(df,sex == 'female'), aes(fill=sex), alpha = 0.2) +
    scale_fill_manual(name="sex", values=c("blue","red"),labels=c("male","female"))
}

# Create plot to demonstrate that it works
by_sex('smoking')
```

_list some observations from your histograms here using Markdown code_:

- Males are at a higher risk of death by smoking than females by almost 50%.


## Bivariate analysis
As a final step, create a **scatterplot matrix** of all of your risks to visually assess the correlation across death rates.

```{r}
pairs(data=risk.data, ~alcohol_use+drug_use+high_meat+low_exercise+smoking)
```

_list some observations from your histograms here using Markdown code_:

- There seems to be a strong correlation between alcohol use and drug use.

When you've finished this step, compile your code to an `html` document. 